---
title: "p8105_hw6_jy2944"
author: "Jie Yu"
date: "11/25/2018"
output: 
  github_document:
    toc: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(mgcv)


knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%",
  warning = FALSE,
  message = FALSE
)
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Problem 1

### Import and tidy data

Read data through a [GitHub repository](https://github.com/washingtonpost/data-homicides). Create a `city_state` variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. Modifiy `victim_race` to have categories `white` and `non-white`, with `white` as the reference category. Be sure that `victim_age` is numeric.

```{r p1_tidy}
# Read data through a Github repo
homicide_data = read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") %>% 
  mutate(
    # Create a `city_state` variable (e.g. “Baltimore, MD”)
    city_state = str_c(city, ",", state),
    # Create a binary variable indicating whether the homicide is solved: 1 - solved; 0 - unsolved
    resolved = as.numeric(disposition == "Closed by arrest")
    ) %>% 
  # Omit some cities
  filter(!city_state %in% c("Dallas,TX","Phoenix,AZ", "Kansas City,MO", "Tulsa,AL")) %>%
  mutate(
    # Modifiy `victim_race` to have categories `white` and `non-white`, with `white` as the reference
    victim_race = fct_relevel(ifelse(victim_race == "White", "white", "non-white"), "white"),
    # Be sure that `victim_age` is numeric
    victim_age = as.numeric(victim_age)
    )
```

### Fit a logistic regression for one city

For the city of Baltimore, MD, use the `glm` function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race (as just defined) as predictors. Save the output of `glm` as an R object; apply the `broom::tidy` to this object; and obtain the estimate and CI of the **adjusted odds ratio** for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

```{r p1_balti}
balti_logistic = homicide_data %>% 
  # filter the city
  filter(city_state == "Baltimore,MD") %>% 
  # Fit a logistic regression: `resolved` as outcome
  glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) 

balti_logistic %>% 
  broom::tidy() %>% 
  # Obtain odds ratio and its CI
  # Note: logistic model estimates are log odds ratios, so we need to tranform them back
  mutate(
    OR = exp(estimate),
    conf.lower = exp(estimate - std.error * 1.96),
    conf.upper = exp(estimate + std.error * 1.96)
    ) %>%
  select(term, OR, conf.lower, conf.upper, p.value) %>% 
  knitr::kable(digits = 3)
```

The odds ratio for solving solving homicides comparing non-white victims to white victims is 0.441 and the CI is (0.312, 0.620) with 95% confidence interval. It means that homicides in which the victim was non-white were substantially less likely to be resolved than those in which the victim was white (OR = 0.441 < 1, p-value < 0.001). The non-white victims had 0.441 times the odds of having their cases being resolved compared to the white victims. The true odd ratio is between 0.312 and 0.620.

### Fit logistic regression for each of the cities

Now run `glm` for each of the cities in the dataset, and extract the adjusted odds ratio and CI for solving homicides comparing non-white victims to white victims. Do this within a “tidy” pipeline, making use of `purrr::map`, list columns, and `unnest` as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r p1_each_city}
# Contruct a function `or_and_ci` to calculate OR and CI using `glm` results
or_and_ci = function(df){
    glm = glm(resolved ~ victim_age + victim_sex + victim_race, data = df, family = binomial())
    
    logistic_tidy = glm %>% 
      broom::tidy() %>% 
      mutate(
        OR = exp(estimate),
        conf.lower = exp(estimate - std.error * 1.96),
        conf.upper = exp(estimate + std.error * 1.96)
        ) %>%
      select(term, OR, conf.lower, conf.upper, p.value) %>% 
      # filter the term of non-white victims
      filter(term == "victim_racenon-white") %>% 
      select(-term)
    }

city_logistic = homicide_data %>% 
  # nest the data by each city
  group_by(city_state) %>% 
  nest() %>% 
  # Apply the function `or_and_ci` to each city
  mutate(or_and_ci = map(.x = data, ~or_and_ci(.x))) %>% 
  select(-data) %>% 
  unnest()

city_logistic
```




